--- 
title: "Class05" 
format: pdf 
author: "Liz" 
toc: true
---




#Our first plot 

R has base graphics 

```{r}
plot(cars)
```
How would I plot this with 'ggplot2'?
NO!

We need to install all load the ggplot package first.
To install any package in R we use the 
'install.packages()' function
```{r}
library(ggplot2)
```



```{r}
ggplot(cars)
```


Every ggplot needs at least 3 layers:
-**Data** (i.e. the data.frame we have),
-**Aes** (the aesthetic mapping of our data to what we want to plot)
**Geoms** (How we want to plot this stuff!)


```{r}
ggplot(data = cars) + aes(x = speed, y = dist) + 
  geom_point() + geom_smooth(se = FALSE)
```


```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

```{r}
ggplot(data = genes) + aes(x = Condition1, y = Condition2, col = State) + 
  geom_point() 
```

```{r}
nrow(genes)
```

```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
```

```{r}
p
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red") )
```

  
library(gapminder)


```{r}
library(gapminder)
```

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"
```

```{r}
gapminder <- read.delim(url)
```


```{r}
library(dplyr)
```
```{r}
gapminder_2007 <- gapminder %>% filter(year==2007)
```


```{r}
library("gapminder")
```

```{r}
library(dplyr)
```


```{r}
gapminder_2007 <- gapminder %>% filter(year==2007)
```
```{r}
library(ggplot2)
```



```{r}
ggplot(gapminder_2007)
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 10)
```

```{r}
library("gapminder")
```

```{r}
gapminder_top5 <- gapminder
```

```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop))
```


```{r}
ggplot(gapminder_top5) + 
geom_col(aes(x = country, y = pop))
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = continent))
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = lifeExp))
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=country) +
  geom_col(col="gray30") +
  guides(fill="none")
```
```{r}
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()
```
```{r}
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment(aes(x=State, 
                   xend=State, 
                   y=0, 
                   yend=Murder), color="blue") +
  coord_flip()
```


